Titeln på avhandlingens språk

### 2023-02-11 body/part1/overview/methodology.tex
expeca

\subsubsection{Testbeds: Ainur}

>>>>

\subsubsection{ExPECA Testbed \& the Ainur software stack}

After our initial implementation and experimentation with our methodology for \gls{WCA} applications, it became clear that our research would benefit greatly from the establishment of a testbed on which our methodology could be deployed.
A testbed would
\begin{inlineenum}
    \item speed up subsequent research by removing the need for ad-hoc implementations for each new study
    \item provide a standardized baseline, allowing for easier comparisons between studies.
    \todo[inline]{More}
\end{inlineenum}.
Additionally, such a testbed would benefit other research happening in the group.

These reasons led to the establishment of the \gls{EXPECA} testbed at \pgls{KTH}.
\gls{EXPECA} is an \gls{SSF}-funded infrastructure project targeting the development and provisioning of an edge computing infrastructure for research into novel applications and network architectures.
It consists of a cluster of hardware-reconfigurable \gls{COTS} computing nodes interconnected using managed switches and \glspl{SDR}.
This allows us to quickly, on-the-fly, and in an automated fashion change the characteristics of the cluster and the network, in order to study different Edge- and Cloud-computing deployments and the applications that run on them.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{publications/2022Ainur/figures/network}
    \caption{
        Prototype architecture of \acs{EXPECA}.
        Figure originally published in \cref{paper:olguinmunoz2022ainur}.
    }\label{fig:expeca}
\end{figure}

The initial prototype architecture of \gls{EXPECA} is illustrated in \cref{fig:expeca}.
\emph{Workload hosts} act as client-side user devices.
These devices connect to either a cloudlet or cloud computing instances over the \emph{workload data network}.
Through intelligent configuration of \acsp{VLAN} and routing, traffic on this network can be forced to flow through the attached \acsp{SDR}, on of them acting as a wireless connection client and the other as an access point.
This allows us to seamlessly switch out the physical layer for experiments.
Alternatively, workload hosts can connect wirelessly via Wi-Fi to an \acs{SDR} acting as an \acs{AP}.
Finally, the \emph{control network} is used for the out-of-band configuration of testbed resources, and the \acs{VPN} links allow direct tunneled connections to cloud compute nodes on \acs{AWS}.

An implementation of this prototype was employed for the studies in \cref{paper:olguinmunoz2022cleave,paper:olguinmunoz2023realistic}.
As an initial deployment of our testbed concept, it employed
\begin{inlineenum}[itemjoin={{; }}, itemjoin*={{; and finally }}]
    \item \num{10} Raspberry Pi 4 Model B boards, acting as workload hosts
    \item two additional Raspberry Pi 4 Model boards acting as \gls{VPN} gateways as well as hosting necessary time synchronization and \gls{DNS} nameserver software
    \item a cloudlet
    \item a logging and management server
    \item a configurable number of \gls{AWS} \gls{EC2} cloud instances acting as cloud compute nodes
    \item two managed switches,
    \item two \glspl{SDR} with associated radio stack processing hosts
\end{inlineenum}.

The above prototype was formally introduced in \cref{paper:olguinmunoz2022ainur}, together with the software framework employed for its automation and experiment orchestration.
